### 2010-03-24 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: fix substExpr (\substH) type to fit the Agda model. I tried to come up with a nicer type signature, not sure it's a success tho.
Desc@ICFP: beginning of clarification of the indexed functor situation.

\begin{wstructure}
<- Adding hindx have introduced some duplication
    <- indx == hindx 1
    -> We can factor out commonalities 
        /> Obtain an equivalent presentation
        /> Still embeddable (refer to the Agda model)
\end{wstructure}

@@

In the previous section, we have presented the $\Desc$ universe as the
grammar of functors in the category $\Set$. We have seen how to encode
inductive types in this setting. To encode an inductive family indexed
by $\Bhab{\V{I}}{\Set}$, we rely on functors in the category
$\Set^I$. We call these \emph{indexed functors}. 
\note{pwm: What's our story \emph{vis-\`{a}-vis} the 2 notions of indexed
functor at play here: $(I\to\Set)\to\Set$ and $(I\to\Set)\to(I\to\Set)$?}  
In the following, we
implement $I \To \IDesc{I}$ as a grammar for indexed functors:

@@

This generalisation can be achieved by a minor evolution of $\Desc$
(Fig.~\ref{fig:type-former-desc}). Indeed, the functorial nature of the
atomic $\Desc$ is entirely captured by the $\DId$ code. We get an indexed
functor by turning $\DId$ into an index-dependent $\DVar$ code:

@@

The resulting code of $\IDesc$ is presented in
Figure~\ref{fig:idesc}.

@@

\newcommand{\substH}{\F{substExpr}}
\newcommand{\domNat}{dom_{\CN{nat}}}
\newcommand{\domBool}{dom_{\CN{bool}}}
\newcommand{\gammaNat}{\gamma_{\CN{nat}}}
\newcommand{\gammaBool}{\gamma_{\CN{bool}}}

\[
\begin{array}{@{}ll}
\substH  : & \PITEL{\V{\domNat}}{\EnumU} 
             \PITEL{\V{\gammaNat}}{\spi{\domNat}{(\LAM{\_} \IMu{\Ty}{\toIDesc{\HExprVarD{\domNat}}}{\Enat})}} \\
           & \PITEL{\V{\domBool}}{\EnumU} 
             \PITEL{\V{\gammaBool}}{\spi{\domBool}{(\LAM{\_} \IMu{\Ty}{\toIDesc{\HExprVarD{\domBool}}}{\Ebool})}} \\
          &  \begin{array}{@{}ll}
             \PI{\V{\sigma}}{& \PITEL{\V{\domNat}}{\EnumU} \\
                             & \PITEL{\V{\gammaNat}}
                                     {\spi{\domNat}{(\LAM{\_} \IMu{\Ty}{\toIDesc{\HExprVarD{\domNat}}}{\Enat})}} \\
                             & \PITEL{\V{\domBool}}{\EnumU} \\
                             & \PITEL{\V{\gammaBool}}
                                     {\spi{\domBool}{(\LAM{\_} \IMu{\Ty}{\toIDesc{\HExprVarD{\domBool}}}{\Ebool})}} \\
                             & \PI{\V{ty}}{\Ty} (\SUM{\Val{ty}}{\Var{dom_{ty}}{ty}}) \To
                               \IMu{\Ty}{(\toIDesc{\HExprD})}{ty}}
             \end{array}\\
          & \PI{\V{ty}}{\Ty}
            \IMu{\Ty}{\toIDesc{\HExprVarD{dom_{ty}}}}{\V{ty}} \To
            \IMu{\Ty}{\toIDesc{\HExprD}}{\V{ty}}
\end{array}
\]

Where $dom_{ty}$ is short for
 $\case{ty}{\Enat  \To \domNat
         \\ \Ebool \To \domBool}$

@@

\[
\sigma\: \domNat\: \gammaNat\: \domBool\: \gammaBool\: ty\: var \mapsto 
    \discharge\: ty\: dom_{ty}\: \gamma_{ty}\: var 
\]

@@

\[\stk{
\substH\: \domNat\: \domBool\: 
            \gammaNat\: \gammaBool\:
            \sigma\: 
            ty\: 
            term \mapsto  \\
\;\;\ \begin{array}{ll}
       \substI\: & \Ty\: 
                  (\SUM{\Val{}}{\Var{dom_{ty}}{}})\: 
                  \Val{}\:
                  \HExprFreeD\:  \\
                &
                  (\sigma\: \domNat\: \domBool\: \gammaNat\: \gammaBool)\:
                  ty\:
                  term
      \end{array}
}\]

>>>>

\begin{wstructure}
[Outdated: type former presentation instead]
<- Adding hindx have introduced some duplication
    <- indx == hindx 1
    -> We can factor out commonalities 
        /> Obtain an equivalent presentation
        /> Still embeddable (refer to the Agda model)
\end{wstructure}

@@

In the previous section, we have presented the $\Desc$ universe as the
grammar of functors in the category $\Set$. We have seen how to encode
inductive types in this setting. To encode an inductive family indexed
by $\Bhab{\V{I}}{\Set}$, we rely on endofunctors on the category
$\Set^I$. We call these \emph{indexed functors}. In the following, $I
\To \IDesc{I}$ is a grammar for indexed functors: \note{pwm: What's
  our story \emph{vis-\`{a}-vis} the 2 notions of indexed functor at
  play here: $(I\to\Set)\to\Set$ and $(I\to\Set)\to(I\to\Set)$?}

@@

This generalisation can be achieved by a minor evolution of $\Desc$
(Fig.~\ref{fig:type-former-desc}). Indeed, the functorial nature of
$\Desc$ is entirely captured by the $\DId$ code. We get an indexed
functor by turning $\DId$ into an index-dependent $\DVar$ code:

@@

The resulting code of $\IDesc$ is presented in
Figure~\ref{fig:idesc}. A fix-point consists of the fix-point of the
\emph{family} of functors as a whole. Hence the following definition:

\[\stkl{
\Rule{\stk{ 
          \Gamma \vdash \Bhab{I}{\Set} \\
          \Gamma \vdash \Bhab{R}{I \To \IDesc{I}} \\
          \Gamma \vdash \Bhab{i}{I}}}
     {\Gamma \vdash \Bhab{\IMu{I}{R}{i}}{\Set}} \qquad
\Rule{\stk{
          \Gamma \vdash \Bhab{I}{\Set} \\
          \Gamma \vdash \Bhab{R}{I \To \IDesc{I}} \\
          \Gamma \vdash \Bhab{i}{I} \\
          \Gamma \vdash \Bhab{x}{\idescop{R\: i}{I}{(\IMu{I}{R}{})}}}}
     {\Gamma \vdash \Bhab{\Con{x}}{\IMu{I}{R}{i}}}
}\]

@@

\newcommand{\substH}{\F{substExpr}}
\newcommand{\domNat}{dom_{\CN{nat}}}
\newcommand{\domBool}{dom_{\CN{bool}}}
\newcommand{\gammaNat}{\gamma_{\CN{nat}}}
\newcommand{\gammaBool}{\gamma_{\CN{bool}}}
\newcommand{\GammaSpi}[2]{\Gamma_{\CN{ty}}~#1~#2}

\[
\begin{array}{@{}ll}
\substH  : & \PITEL{\V{\dom}}{\EnumU} \\
           & \PITEL{\V{\gammaNat}}{\GammaSpi{dom}{\Enat}} 
             \PITEL{\V{\gammaBool}}{\GammaSpi{dom}{\Ebool}} \\
           & \begin{array}{@{}ll}
             \PI{\V{\sigma}}{& \PITEL{\V{dom}}{\EnumU} \\
                             & \PITEL{\V{\gammaNat}}
                                     {\GammaSpi{dom}{\Enat}} \\
                             & \PITEL{\V{\gammaBool}}
                                     {\GammaSpi{dom}{\Ebool}} \\
                             & \PI{\V{ty}}{\Ty} (\SUM{\Val{ty}}{\Var{dom}{ty}}) \To \\
                             & \IMu{\Ty}{(\toIDesc{\HExprD})}{ty}}
             \end{array}\\
          & \PI{\V{ty}}{\Ty}
            \IMu{\Ty}{\toIDesc{\HExprVarD{dom}}}{\V{ty}} \To \\
          & \IMu{\Ty}{\toIDesc{\HExprD}}{\V{ty}}
\end{array}
\]

Where $\GammaSpi$ corresponds to a context, defined by:

\[\stk{
\GammaSpi{}{} : \EnumU \To \Ty \To \Set                                             \\
\GammaSpi{dom}{ty} = \spi{dom}{(\LAM{\_} \IMu{\Ty}{\toIDesc{\HExprVarD{dom}}}{ty})} 
}\]

@@

\[
\sigma\: dom\: \gammaNat\: \gammaBool\: ty\: var \mapsto 
    \discharge\: ty\: dom\: \gamma_{ty}\: var 
\]

Where $\gamma_{ty}$ is short for
 $\caseB{ty}{\Enat  \To \gammaNat \\
             \Ebool \To \gammaBool}$

@@

\[\stk{
\substH\: dom\:
          \gammaNat\: \gammaBool\:
          \sigma\: 
          ty\: 
          term \mapsto  \\
\;\;\ \begin{array}{ll}
       \substI\: & \Ty\: 
                  (\SUM{\Val{}}{\Var{dom}{}})\: 
                  \Val{}\:
                  \HExprFreeD\:  \\
                &
                  (\sigma\: dom\: \gammaNat\: \gammaBool)\:
                  ty\:
                  term
      \end{array}
}\]
